Long term cloud enhancement events of global solar irradiance over Thessaloniki, Greece

### 2024-06-09 article/article.Rmd
Commit 2024-06-09 18:17

For the detection of CE cases we established a baseline of irradiance above which we
characterized each data point as an enhancement event and calculated the over
irradiance (OIR). The OIR is defined as the irradiance difference of the measured
one-minute $\text{GHI}_i$ from the $\text{GHI}_\text{ref}$ corresponding to
cloud-free atmosphere. 
<!-- CE identification criterion ($\text{OIR}_i =
\text{GHI}_i - \text{GHI}_\text{CSlim}$), as defined in
Equation\nobreakspace\ref{eq:CE4}. -->
First, we tested two simple approaches for the determination of
$\text{GHI}_\text{ref}$: The Haurwitz's model [@Haurwitz1945], which is a simple
clear sky model and was already adjusted and applied to our data [@Natsis2023], and
the total solar irradiance (TSI) at the top of the atmosphere, adjusted for the
Sun-Earth distance.  We have tested both methods by using an appropriate relative
threshold and/or an additional constant offset.  The initial results showed that we
can detect a big portion of the CE events. However, by inspecting the daily plots of
irradiance it became evident that changes on the atmospheric conditions introduced
numerous false positive and false negative results.  To produce a more accurate
reference, we had to take into account more factors that effect the clear sky
radiation. So we used a radiation transfer model in order to include the effects of
ADO and the water vapors.

## Modeled clear Sky Irradiance

@@

We approximated the expected clear sky GHI by using the Libradtran radiation transfer
model [@Emde2016], a similar approach, was also used by @Vamvakas2020 for creating a
clear sky reference for studding cloud enhancement events.  Main factors responsible
for the attenuation of the broadband downward solar radiation in the atmosphere are
the aerosols and the water vapors.  Because of the lack of observational data for the
whole period, we used some long term climatological data to recreate the whole
period.  Fortunately, our site participates in the Aerosol Robotic Network (AERONET)
[@Giles2019; @Buis1998], as there is in operation a Cimel photometer since 2003,
collocated with the CM-21 pyranometer.  The mean monthly AOD on different wavelengths
are provided by AERONET, along with the equivalent water column height in the
atmosphere.

<!-- Libradtran options -->
<!-- 
For completeness, we will describe here the main points of the radiation simulation
procedure. We used as input the spectrum of @Kurucz1994 in the range $280$ to
$2500\,\text{nm}$, with the Libradtran radiation transfer solver "disort" on a
"pseudospherical" geometry and the "LOWTRAN" gas parameterization. For each
combination of conditions we use a SZA step of $0.2^\circ$.
<!-- Atmospheric options -->
For the atmospheric characteristics, we iterated on combinations of AOD at
$500\,\text{nm}$ ($\tau_{500\text{nm}}$) with additional offsets of $\pm1$ and
$\pm2\sigma$, and water column ($w$), also with offsets of $\pm1$ and $\pm2\sigma$.
We applied them on two atmospheric profiles, from the Air Force Geophysics
Laboratory (AFGL). The "AFGL atmospheric constituent profile, midlatitude summer"
(afglms) and the "AFGL atmospheric constituent profile, midlatitude winter" (afglmw)
[@Anderson1986].
-->

>>>>

For the detection of CE cases we established a baseline of irradiance above which we
characterized each data point as an enhancement event and calculated the over
irradiance (OIR). The OIR is defined as the irradiance difference of the measured
one-minute $\text{GHI}_i$ from the $\text{GHI}_\text{ref}$ corresponding to
cloud-free atmosphere. 
<!-- CE identification criterion ($\text{OIR}_i =
\text{GHI}_i - \text{GHI}_\text{CSlim}$), as defined in
Equation\nobreakspace\ref{eq:CE4}. -->
First, we tested two simple approaches for the determination of
$\text{GHI}_\text{ref}$: The Haurwitz's model [@Haurwitz1945], which is a simple
clear sky model and was already adjusted and applied to our data [@Natsis2023], and
the total solar irradiance (TSI) at the top of the atmosphere, adjusted for the
Sun-Earth distance.  We have tested both methods by using an appropriate relative
threshold and/or an additional constant offset.  The initial results showed that we
can detect a big portion of the CE events. However, by inspecting the daily plots of
irradiance it became evident that changes in atmospheric conditions introduced
numerous false positive and false negative results.  The main reason for these
discrepancies is the variability of the effects of aerosols and water vapor which
were not taken into account in the two simple methods.  To produce a more
representative reference  we included the effects of these factors through variables
using a radiative transfer model (RTM).  The applied methodology is discussed in
section\nobreakspace{}\ref{rtmcs}

## Modeled clear Sky Irradiance {#rtmcs}

@@

<!-- Libradtran options -->
We approximated the expected clear sky $\text{GHI}_\text{ref}$ with the radiative
transfer model uvspec, part of libRadtran [@Emde2016], similarly to the approach used
by @Vamvakas2020.  In uvspec we used the solar spectrum of @Kurucz1994 in the range
$280$ to $2500\,\text{nm}$, the radiative trasfer solver "disort" in
"pseudospherical" geometry and the "LOWTRAN" gas parameterization.  The model was run
for a range of variables in order to create a look up table (LUT) for the estimation
of the cloud-free reference irradiance for each individual observation of our
dataset. In this context, the model was run for SZAs in the range $10$ -- $90^\circ$
with a step of $0.2^\circ$ and for the atmospheric profiles of the Air Force
Geophysics Laboratory [@Anderson1986] midlatitude summer and midlatitude winter,
representative of the warm and cold seasons.

Main factors responsible for the attenuation of the broadband downward solar
radiation under cloud free atmospheres are aerosols and water vapor.  At
Thessaloniki, such measurements are available since 2003 from the Cimel sun
photometer that belongs to the Aerosol Robotic Network (AERONET) [@Giles2019;
@Buis1998].  The mean monthly AOD on different wavelengths are provided by
AERONET, along with the equivalent water column height in the atmosphere.

<!-- Atmospheric options -->
For the atmospheric characteristics, we iterated on combinations of AOD at
$500\,\text{nm}$ ($\tau_{500\text{nm}}$) with additional offsets of $\pm1$ and
$\pm2\sigma$, and water column ($w$), also with offsets of $\pm1$ and $\pm2\sigma$.
